---
title: "HW10_Markdown"
output:
  pdf_document: default
  html_document: default
date: "2024-04-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a

My partner is Heinrick Gonzalez. Our research question is "How do different 
factors affect tourism?"

## b

Heinrick is the only other group member for this project. The grade I would
give him is an A because he has been really responsive and helpful with
collecting data, analysis, and overall research.

## c

I looked from other papers on google scholar and they said that transportation,
price of hotels and commodities, education, hospital facilities and languages
can all affect tourism.

## d

We found 3 datasets. One had all the general information about countries like
the gdp, population, land size, etc. The second dataset had information
relating to happiness so it measured social support, life expectency, freedom,
etc. The third dataset provided the tourism information about how many people
come into a country in a given year.
## e

```{r e}
combined = read.csv("~/Downloads/combined.csv")
dim(combined)
summary(combined)
```

There are 50 total variables and 145 observations. Each observation representats
a different country. Some countries were ommitted because the tourism data was 
not collected for them. 19 of the variables are doubles and the remainder are
characters. However, some that are characters should be integers or doubles
which we will have to change in the future.

## f

One method we plan on using is prediction, specifically linear regressions. We
are doing this just to see if we can find high correlations among certain
variables. We will also plan on doing unsupervised learning to see if there
are clusters we can identify.
