---
title: "HW11_Markdown"
output: pdf_document
date: "2024-04-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Manay Divatia
# md46245

## a
```{r a}
load("~/Downloads/newspapers.RData")
library(ggplot2)
str(papers)

ggplot(papers, aes(x = nslant)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  geom_vline(aes(xintercept = median(nslant)), color = "red", linetype = "dashed") +
  labs(title = "Distribution of Ideological Slant in Newspapers",
       x = "Ideological Slant (nslant)",
       y = "Frequency") +
  theme_minimal()

left_most_slant <- papers[which.min(papers$nslant), ]
right_most_slant <- papers[which.max(papers$nslant), ]

cat("Newspaper with the largest left-wing slant:", left_most_slant$paper, "\n")
cat("Newspaper with the largest right-wing slant:", right_most_slant$paper, "\n")
```
The newspaper with the largest left-wing slant was Chicago Defender. The
newspaper with the largest right-wing slant was The Daily Sentinel.


## b
```{r b}
library(wordcloud)
terms <- dtm$dimnames$Terms
#freq <- colSums(as.matrix())
#dtm_df <- data.frame(terms = terms, freq = dtm$ncol)
#dtm_df <- dtm_df[order(-dtm_df$freq), ]

#wordcloud(words = dtm_df$terms, freq = dtm_df$freq, max.words = 20)
#left_most_newspapers <- papers[order(papers$nslant)][1:round(nrow(papers)/10), ]
#right_most_newspapers <- papers[order(papers$nslant, decreasing = TRUE)][1:round(nrow(papers)/10), ]
#dtm_df <- as.data.frame(dtm)

#left_most_dtm <- dtm_df[rownames(dtm_df) %in% left_most_newspapers$newsid, ]
#right_most_dtm <- dtm_df[rownames(dtm_df) %in% right_most_newspapers$newsid, ]

#wordcloud(words = rownames(left_most_dtm), freq = colSums(left_most_dtm), max.words = 20)
#wordcloud(words = rownames(right_most_dtm), freq = colSums(right_most_dtm), max.words = 20)
```
A lot of the words differ because they talk about different things. However,
the general tone of the topics are more or less the same, it is just the
specific words that are completely different.

## c
```{r c}
house_avg_slant <- aggregate(cslant ~ state, data = subset(cong, chamber == "H"), FUN = mean)
senate_avg_slant <- aggregate(cslant ~ state, data = subset(cong, chamber == "S"), FUN = mean)
newspaper_avg_slant <- aggregate(nslant ~ state, data = papers, FUN = mean)

ggplot() +
  geom_point(data = house_avg_slant, aes(x = cslant, y = newspaper_avg_slant$nslant[2:51]), color = "green") +
  geom_smooth(data = house_avg_slant, aes(x = cslant, y = newspaper_avg_slant$nslant[2:51]), method = "lm", color = "red") +
  labs(title = "Comparison of House Congressional Slant and Newspaper Slant by State",
       x = "House Congressional Slant",
       y = "Newspaper Slant") +
  theme_minimal()

ggplot() +
  geom_point(data = senate_avg_slant, aes(x = cslant, y = newspaper_avg_slant$nslant[2:51]), color = "green") +
  geom_smooth(data = senate_avg_slant, aes(x = cslant, y = newspaper_avg_slant$nslant[2:51]), method = "lm", color = "red") +
  labs(title = "Comparison of Senate Congressional Slant and Newspaper Slant by State",
       x = "Senate Congressional Slant",
       y = "Newspaper Slant") +
  theme_minimal()
```

From the plots, I don't think we can make any conclusive claims about whether
or not the newspapers are directly influenced by the political language of
elected officials. While, the regression shows a positive correlation we can
see a large variance band which tells us how unsure we are about the regression
line.

## d
```{r d}
cong_tx <- subset(cong, state == "TX")
papers_tx <- subset(papers, state == "TX")
merged_tx <- merge(cong_tx, papers_tx, by = c("district", "state"))
house_avg_slant_tx <- aggregate(cslant ~ district, data = subset(merged_tx, chamber == "H"), FUN = mean)
newspaper_avg_slant_tx <- aggregate(nslant ~ district, data = merged_tx, FUN = mean)

ggplot() +
  geom_point(data = house_avg_slant_tx, aes(x = cslant, y = newspaper_avg_slant_tx$nslant), color = "green") +
  geom_smooth(data = house_avg_slant_tx, aes(x = cslant, y = newspaper_avg_slant_tx$nslant), method = "lm", color = "red") +
  labs(title = "Comparison of House Congressional Slant and Newspaper Slant by District in Texas",
       x = "House Congressional Slant",
       y = "Newspaper Slant") +
  theme_minimal()
```

These results tell a different story but it isn't a conclusive story. The
variance is very high so I think we can't make any definitive conclusions based
on the results of this graph.

## e
```{r e}

#library(tm)
#dtm_tdm <- TermDocumentMatrix(dtm)

#tfidf <- weightTfIdf(dtm_tdm)
#tfidf_df <- as.data.frame(as.matrix(tfidf))

#home_news_tribune <- tfidf_df[rownames(tfidf_df) == "Home News Tribune", ]

#top_terms <- sort(home_news_tribune, decreasing = TRUE)[1:10]
#print(top_terms)
```


