---
title: "HW3_Markdown"
output: pdf_document
date: "2024-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Manay Divatia
# md46245
# 2/14/24

# Problem 1

## a
```{r a}
gay_df = read.csv("~/Downloads/gay.csv")
wave1_df = subset(gay_df, subset = wave == 1)
summary(wave1_df)
no_contact_1 = subset(wave1_df, subset = (treatment == "No Contact"))
dim(no_contact_1)
marriage_gay_1 = subset(wave1_df, subset = (treatment == "Same-Sex Marriage 
                                            Script by Gay Canvasser"))
dim(marriage_gay_1)
marriage_straight_1 = subset(wave1_df, subset = (treatment == "Same-Sex Marriage
                                                 Script by Straight Canvasser"))
dim(marriage_straight_1)
```
We see a lot more instances of no contact than with any other treatment which 
may suggest that it wasn't random after all. Another thing to note is that the 
treatment of Same-Sex Marriage Script by Gay Canvasser seemed to be almost 
double of the other treatments aside from the No Contact treatment. 
This also may indicate that the study was not random.
## b
```{r b}
mean(subset(gay_df, subset=(wave == 2 & study == 1 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 2 & study == 1 & treatment == "Same-Sex 
                            Marriage Script by Straight Canvasser"))$ssm)
```
From this, we can see that the average support for study 1 where there were gay
canvassers was 3.13 and for that of straight canvassers it was 3.16. This is not
really a large difference to indicate anything with full certainty.
## c
```{r c}
mean(subset(gay_df, subset=(wave == 2 & study == 1 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 2 & study == 1 & treatment == "Recycling 
                            Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 2 & study == 1 & treatment == "Same-Sex 
                            Marriage Script by Straight Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 2 & study == 1 & treatment == "Recycling 
                            Script by Straight Canvasser"))$ssm)
```
I think the purpose of having authors use a script to encourage people to recycle
helped to show some inherent biases people may have even when hearing about
nothing related to same sex marriage. When comaparing the results of the gay 
canvassers for the same-sex marriage and recycling scripts, we found that the
same-sex marriage script had a 3.13 support while the recycling had 3.1 
which isn't enough of a difference to make any claims. When comaparing the 
results of the straight canvassers for the same-sex marriage and recycling 
scripts, we found that the same-sex marriage script had a 3.16 support while 
the recycling had 3 which is slightly larger and may be indicative of some trend.
## d
```{r d}
mean(subset(gay_df, subset=(wave == 7 & study == 1 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 7 & study == 1 & treatment == "Same-Sex 
                            Marriage Script by Straight Canvasser"))$ssm)
```
It seems that there could be some lasting effects. After computing the mean, 
we find that with the gay canvasser, the average was 3.37 while with the straight
canvasser, it was 3.27. This difference is about .1 which is something that
could or couldn't be attributed to the canvassers.
## e
```{r e}
no_contact_2 = subset(gay_df, subset = (wave == 1 & study == 2 & treatment == 
                                          "No Contact"))
dim(no_contact_2)
marriage_gay_2 = subset(gay_df, subset = (wave == 1 & study == 2 & 
                                            treatment == "Same-Sex Marriage 
                                          Script by Gay Canvasser"))
dim(marriage_gay_2)
```
It does seem that there is some randomization because the subsets seems to be
around the same size. Probably what is more important is that there isn't a 
drastic difference.
## f
```{r f}
mean(subset(gay_df, subset=(wave == 2 & study == 2 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
```
Yes it looks to be consistent to some extent. The data showed in the first study
a ssm of 3.13 while in the second study, a ssm of 3.11. If they were properly
randomized, they wouldn't be too different in score.
## g
```{r g}
mean(subset(gay_df, subset=(wave == 1 & study == 2 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 2 & study == 2 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 3 & study == 2 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 4 & study == 2 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
mean(subset(gay_df, subset=(wave == 7 & study == 2 & treatment == "Same-Sex 
                            Marriage Script by Gay Canvasser"))$ssm)
```
The first wave is the lowest with a score of 2.97. The next 3 waves hover 
around 3.1 +- .1. The final wave jumps up to 3.33. Ultimately, it looks like
the score increases throughout the waves in study 2. However, without the 5th 
or 6th wave, we can't know the full story.

# Problem 2

## a
```{r a2}
leaders = read.csv("~/Downloads/leaders.csv")
summary(leaders)
dim(leaders)
unique(leaders$country)
(2001 - 1878) / nrow(leaders)
```
We can see from the dim() method that there are 250 assassination attempts. 
From unique(), we know that there are 88 countries that have experience at least
one leader assassination attempt. From the summary function, we found that this
data ranged from 1878 to 2001. Finding mean assassinations per year, we get .492
which could be representated as about 1 assassination attempt every 2 years.
## b
```{r b2}
leaders$dead <- ifelse(leaders$result == "dies within a day after the attack" | 
                         leaders$result == "dies between a day and a week" | 
                         leaders$result == "dies between a week and a month" | 
                         leaders$result == "dies, timing unknown", 1, 0)

```
It does speak to the validity of the assumption that attempts are randomly
determined because there seems to be a distribution of ways and counts in which
a leader dies or doesn't die.
## c
```{r c2}
mean(leaders$politybefore)
mean(leaders$polityafter)
mean(leaders$age[leaders$dead == 0])
mean(leaders$age[leaders$dead == 1])
```
The polity before is -1.51 and after is -1.65. This shows a movement towards
hereditary monarchy as defined by the Polity Project. As for age, the average
age for leaders who survived was 52.7 while the average age for leaders who died
was 56.5 which means that older leaders were maybe more likely to die.
## d
```{r d2}
mean(leaders$interwarbefore)
mean(leaders$interwarafter)
mean(leaders$civilwarbefore)
mean(leaders$civilwarafter)
leaders$warbefore <- ifelse(leaders$civilwarbefore == 1 | 
                              leaders$interwarbefore == 1, 1, 0)
mean(leaders$warbefore)
```
In terms of international and civil war, both dropped in percentage from before
to after. For international, it went from .188 to .148 and for civil war, 
it went from .216 to .184. After creating the warbefore column and looking
at the mean, we find that 36.8% of assassinations had a civil or international 
war happen within 3 years prior to the assassination attempt.

## e
```{r e2}
leader_dead_df = subset(leaders, subset=(dead == 1))
mean(leader_dead_df$polityafter - leader_dead_df$politybefore)
leaders$warafter <- ifelse(leaders$civilwarafter == 1 | 
                              leaders$interwarafter == 1, 1, 0)
mean(leaders$warafter[leaders$dead == 1] - leaders$warbefore[leaders$dead == 1])
```
From my findings, it looks like a successful leader assassination actually leads
to more hereditary monarchy. This is because the average difference is -0.5.
Since the figure is negative, it means that after assassinations, countries
tend ot move towards monarchies. From my findings, it looks like that war 
decreases after a successful assasination. This is because my findings show that
war after is lower than war before which means that war went down. Specifically,
the mean was -.148 or -14.8%.