---
title: "HW4_Markdown"
output:
  pdf_document: default
  html_document: default
date: "2024-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Manay Divatia
# md46245
# 1

## a
```{r a, results = "hide"}
rosca = read.csv("~/Downloads/rosca.csv")
rosca$treatment <- ifelse(rosca$encouragement == 1, "control", 
                          ifelse(rosca$safe_box == 1, "safebox", "lockbox"))
nrow(rosca[rosca$treatment == "control", ])  
nrow(rosca[rosca$treatment == "safebox", ])  
nrow(rosca[rosca$treatment == "lockbox", ])
```
There are 111 individuals in the control, 117 for the safebox treatment, and 
195 for the lockbox treatment. 

## b
```{r b, results = "hide"}
nrow(rosca)
rosca_followup <- rosca[rosca$has_followup2 == 1, ]
nrow(rosca_followup)
nrow(rosca_followup[rosca_followup$treatment == "control", ])  
nrow(rosca_followup[rosca_followup$treatment == "safebox", ])  
nrow(rosca_followup[rosca_followup$treatment == "lockbox", ])  
```
There are 393 left in the data which is a lot considering we started with 423.
The most people (11) drop from the lockbox treatment which could skew the data
later on because those people were less likely anyway to invest money.

## c
```{r c, results = "hide"}
mean(rosca_followup[rosca_followup$treatment == "control", ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "safebox", ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "lockbox", ]$fol2_amtinvest)
```
The safebox had by far the most invested with 408. This was followed by the
lockbox which had 308 and the control which had 258. This means that both the
treatments seemed to have a positive effect on the amount invested in the end.

## d
```{r d, results = "hide"}
function_d <- function(treatment) {
  print(mean(rosca_followup[rosca_followup$treatment == treatment, 
                            ]$bg_female))
  print(mean(rosca_followup[rosca_followup$treatment == treatment, 
                            ]$bg_married))
  print(mean(rosca_followup[rosca_followup$treatment == treatment, 
                            ]$bg_b1_age))
}
function_d("control")
function_d("safebox")
function_d("lockbox")
```
The safebox treatment group had the most females and the least married people.
These could affect the results. However, for the rest of the data, the
treatments and control seemed to be similar in terms of percentage female, age,
and marriage status.

## e
```{r e, results = "hide"}
mean(rosca_followup[rosca_followup$treatment == "control" & 
                      rosca_followup$bg_married == 1, ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "safebox" & 
                      rosca_followup$bg_married == 1, ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "lockbox" & 
                      rosca_followup$bg_married == 1, ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "control" & 
                      rosca_followup$bg_married == 0, ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "safebox" & 
                      rosca_followup$bg_married == 0, ]$fol2_amtinvest)
mean(rosca_followup[rosca_followup$treatment == "lockbox" & 
                      rosca_followup$bg_married == 0, ]$fol2_amtinvest)
```
In all three treatments, when the individual was married, the amount invested 
was less. This was the biggest changei n the safebox treatment which went from
an average of 470 to 241. Clearly, marriage status played a large role. It
would be interesting to see how gender and age also affected amount invested.

# 2

## a
```{r 2a, results = "hide"}
predimed = read.csv("~/Downloads/predimed.csv")
total_rows = nrow(predimed)
nrow(predimed[predimed$event == "Yes", ]) / total_rows
nrow(predimed[predimed$group == "MedDiet + Nuts" | 
                predimed$group == "MedDiet + VOO" , ]) / total_rows
nrow(predimed[(predimed$group == "MedDiet + Nuts" | predimed$group == 
                 "MedDiet + VOO") & predimed$event == "Yes", ]) / total_rows
any_meddiet <- predimed[predimed$group == "MedDiet + Nuts" | predimed$group
                        == "MedDiet + VOO" , ]
nrow(any_meddiet[any_meddiet$event == "Yes", ]) / nrow(any_meddiet)
```
P(event) is the chance of the event happening which was 3.98%. P(any MedDiet)
was the chance of the individual being on any of the mediterranean diets which
was 67.7%. P(event, any MedDiet) is the chance of the cardiac event happening
and the individual being on any of the mediterranean diets which was 2.45%. 
P(event | any MedDiet) is the chance of the cardiac event happening given that
the individual is on any of the mediterranean diets which was 3.62%.

## b
```{r 2b, results = "hide"}
control_predimed <- predimed[predimed$group == "Control", ]
nrow(control_predimed[control_predimed$event == "Yes", ]) / 
  nrow(control_predimed)
```
P(event | control) is the chance of a cardiac event given that the individual
is not given the diet. The chance was 4.7%. From the previous question, we see
a change from 3.62 to 4.7% which seems like a big difference to consider that
the mediterranean diet decreases the chance of a cardiac event. However, it
is important to note that there could be other factors involved.

## c
Additional information like eating habits of the individual, prior health
concerns, family history, and location could all affect our results and should
be tracked so that we can consider them.

## d
```{r 2d, results = "hide"}
nrow(any_meddiet[any_meddiet$sex == "Male" & any_meddiet$event == "Yes", ]) / nrow(any_meddiet[any_meddiet$sex == "Male", ])
nrow(any_meddiet[any_meddiet$sex == "Female" & any_meddiet$event == "Yes", ]) / nrow(any_meddiet[any_meddiet$sex == "Female", ])
nrow(predimed[predimed$sex == "Male" & predimed$event == "Yes" & predimed$group
              == "Control", ]) / nrow(predimed[predimed$sex == "Male" & predimed$group=="Control", ])
nrow(predimed[predimed$sex == "Female" & predimed$event == "Yes" & 
                predimed$group == "Control",]) / 
  nrow(predimed[predimed$sex == "Female" & predimed$group == "Control", ])
```
In both cases of control and mediterranean diet, it looks like the likelihood
of a cardiac event is lower for women than it is for men which could be
something to consider when looking at general trends in our data. Additionally,
by splitting it up, it looks like the chance of a cardiac event is higher for
both males and females when given the control.

