---
title: "HW9_Markdown"
output: pdf_document
date: "2024-04-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Manay Divatia
# md46245

```{r a}
library(dplyr)
library(ggplot2)
library(tidyr)
library(cluster)  
social = read.csv("~/Downloads/social_marketing.csv")
summary(social)
```
From the summary, we can get some key metrics. What was most important to me
was seeing if there were any null values. Based on this we know that there
are not any which is good. 

```{r b}
interest_distribution <- colSums(select(social, -X))
print(interest_distribution)
```
From this information, we are able to see how often each category appears. From
this, we can see that chatter, health_nutrition, cooking, photo_sharing,
college_uni, travel, and sports_fandom, and personal_fitness. Immediately, we
can see how some of these go together. For example, personal fitness, health
nutrition, and cooking are all similar topics. Additionally travel and photo
sharing could also go together.

```{r c}
social = select(social, -X)
set.seed(123)
k <- 5 
kmeans_model <- kmeans(social, centers = k)

social_clusters <- social %>%
  mutate(cluster = kmeans_model$cluster)

cluster_summary <- social_clusters %>%
  group_by(cluster) %>%
  summarize_all(mean)

print(cluster_summary)
```



```{r d}
cluster_profiles <- cluster_summary %>%
  gather(key = "interest", value = "mean_count", -cluster)

ggplot(cluster_profiles, aes(x = interest, y = mean_count, fill = factor(cluster))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Cluster Profiles",
       x = "Interest",
       y = "Mean Post Count") +
  theme_minimal()
```


This graph shows the mean post counts for each category based on the 5 
clusters. We actually get good information from this. To start, we can see that
the largest blue line in the middle is in cluster 4 and is the health_nutrition
category. What we can alos see from this is that, unsurprisingly, the cooking
and personal_fitness categories also have high mean post counts for this
clusters which makes sense. We see a similar idea in cluster 5 which is
interesting. Cluster 5 has 2 categories with large mean post counts which are
college_uni and online_gaming. This also makes sense since a lot of college
students do play video games and are likely to tweet about it. 

Ultimately, by clustering and visualizing the data, we see some interesting
trends in the data that we couldn't have seem if we just looked at the numbers.

